Does It Scale? Symbolic Execution of PHP Web Applications

### 2017-01-19 paper.tex
draft

Knowledge about different output variants is leveraged by a number of
analyses for the domain of web applications. Previous work presented tools
for detecting and locating HTML validation errors
\cite{Nguyen:2011:AFH:2190078.2190142}, computing program slices across
server-side and client-page code \cite{Nguyen:2015:CPS:2786805.2786872}, or
easing development and maintenance by extending IDE support for web applications
with code navigation
\cite{Nguyen:2015:VIS:2819009.2819140,Nguyen:2014:BCG:2635868.2635928}.

@@

	\item \emph{Varis}: Editor services such as “jump to declaration” are
	nontrivial for dynamic web applications due to their staged nature. Varis
	provides editor services across stages for client-side code: Starting with an
	ouput approximation a callgraph with conditional edges is constructed, which
	allows navigation in client-site code (HTML, JS and CSS) although it is
	embedded in server-side code
	\cite{Nguyen:2015:VIS:2819009.2819140,Nguyen:2014:BCG:2635868.2635928}.
	\item \textcolor{gray}{Automated Web Application Testing Using Seacrch
	Software Engineering} \cite{Alshahwan2011}
\end{itemize}

@@

As the observations so far did only refer to \sf{WordPress} as a large-scale
system, we wanted to study the impeding impact of dynamic features to symbolic
execution for a wider selection of PHP systems systematically apart from the
systems that \sf{PhpSync} was applied to for two reasons. First, the usage of
dynamic features, not only of includes and function invocations, varies from
system to system \cite{hills2015variable}. Second, we were not able to obtain results for
all entry points possible for \sf{WordPress} and hence could not compare our
approximatio results to those of \sf{PhpSync} where client page output
approximation is evaluated for all entry points. Therefore, we decided to
systematically evaluate our symbolic execution engine for systems apart from
\sf{WordPress}, which is documented in the following Section
\ref{sec:evaluation}.

@@

First, the evaluation of the output approximation using the symbolic execution
engine, \sf{Symex}, in previous work \cite{Nguyen:2014:BCG:2635868.2635928} has
shown high code coverage for small systems. As our symbolic execution
re-implements \sf{Symex} and extends with support for additional language
features, we ask whether we can replicate high coverage results at least on a small-scale. Therefore, our PHP
corpus contains four small-scale systems used in previous work \cite{Nguyen:2014:BCG:2635868.2635928}.

Second, we ask whether system size and usage of modern language features are
factors affecting the practicality of our symbolic execution engine. We
selected both a range of large-scale and small-scale modern PHP systems to
better separate and understand the circumstances under which conceptual
limitations might affect practicality. For the selection of large-scale systems
we borrow three systems from a case study addressing the feature usage in PHP
systems. Some more detailed description of this case study can be found in
Section \ref{sec:related_work}. For the small-scale systems we selected five systems from
a list of recent Content Management Systems (CMS) written in PHP \cite{codegeekz}. For both
selections we were bound to select a smaller number of system than we intended
to since the parser used in our implementation, \sf{Quercus}\footnote{Quercus
footnote}, did not support a number of language features. The full list of PHP
systems for that we evaluate our symbolic execution engine along with
descriptive code metrics can be found in Table \ref{corpus}.

@@

We evaluated this heuristic manually with a sample of 400 string
literals randomly selected from our corpus of PHP systems (see Table
\ref{corpus}). We measured for our heuristic classifier a precision of 94 percent, and a
recall of 50 percent. This means that six percent of the string literals are
classified incorrectly as false positives. In turn, the classifier is highly
distinctive as 96 percent of the string literals are classified correctly. The
recall of 50 percent means that half of the string literals, which we manually 
classified as output candidates, actually were responsive to the classifier. We
attempted to increase recall by looking for further distinctive properties to build a
classifier from, but could not do so without decreasing precision. 
\textcolor{gray}{In spite of missing half of the output candidates, we decided
to use a simple, yet precise and distinctive classifier.}

@@

\paragraph{Dynamic Features}
According to a case study by Hills et al. \cite{Hills:2013:ESP:2483760.2483786}
of the feature usage in PHP, based on a corpus of state-of-the-art-projects,
dynamic includes are less frequently used than static includes, but usage
frequency is varying from system to system. Further work by Hills et al.
\cite{hills2014static,hills2014php} approaches static resolution of dynamic includes using
both context-insensitive resolution on file level by simplifying PHP constants; and context-sensitive on program level for transitive includes. In spite of
promising results, these approaches, similar to our results, face limitations
for truly dynamic includes if resolution is not sound due to information not
being available in the source code like database query results. Something about
dynamic features and resolution.

>>>>

Knowledge about different output variants is leveraged by a number of
analyses for the domain of web applications. Previous work presented tools
for detecting and locating HTML validation errors
\cite{Nguyen:2011:AFH:2190078.2190142}, computing program slices across
server-side and client-page code \cite{Nguyen:2015:CPS:2786805.2786872}, or
easing development and maintenance by extending IDE support for web applications
with code navigation \cite{Nguyen:2015:VIS:2819009.2819140,Nguyen:2014:BCG:2635868.2635928}.
Moreover, knowledge of output variants can increases code coverage for
output-oriented testing \cite{Alshahwan2011}.

@@

	\item \emph{Varis}: Editor services such as “jump to declaration” are
	nontrivial for dynamic web applications due to their staged nature. Varis
	provides editor services across stages for client-side code: Starting with an
	ouput approximation a callgraph with conditional edges is constructed, which
	allows navigation in client-site code (HTML, JS and CSS) although it is
	embedded in server-side code
	\cite{Nguyen:2015:VIS:2819009.2819140,Nguyen:2014:BCG:2635868.2635928}.

\end{itemize}

@@

As the observations so far did only refer to \sf{WordPress} as a large-scale
system, we wanted to study the impeding impact of dynamic features to symbolic
execution for a wider selection of PHP systems systematically apart from the
systems that \sf{PhpSync} was applied to for two reasons.

First, the usage of dynamic features, not only of includes and function invocations, varies from
system to system \cite{hills2015variable}.

Second, we were not able to obtain results for all entry points possible for
\sf{WordPress} and hence could not compare our approximation results to those of
\sf{PhpSync} where client page output approximation is evaluated for all entry points. Therefore, we decided to
systematically evaluate our symbolic execution engine for systems apart from
\sf{WordPress}, which is documented in the following Section
\ref{sec:evaluation}.

@@

First, the evaluation of the output approximation using the symbolic execution
engine, \sf{PhpSync}, in previous work \cite{Nguyen:2014:BCG:2635868.2635928}
has shown high code coverage for small systems. As our symbolic execution
re-implements \sf{PhpSync} and extends with support for additional language
features, we ask: Can  we replicate high coverage results at least on a
small-scale. Therefore, our PHP corpus contains four small-scale systems used
in previous work \cite{Nguyen:2014:BCG:2635868.2635928}.

Second, we ask :Do system size and usage of modern language features
factors affect the practicality of our symbolic execution engine? We
selected both a range of large-scale and small-scale modern PHP systems to
better separate and understand the circumstances under which conceptual
limitations might affect practicality. For the selection of large-scale systems
we borrow three systems from a case study addressing the feature usage in PHP
systems. Some more detailed description of this case study can be found in
Section \ref{sec:related_work}. For the small-scale systems we selected five systems from
a list of recent Content Management Systems (CMS) written in PHP \cite{codegeekz}. For both
selections we were bound to select a smaller number of system than we intended
to since the parser used in our implementation, \sf{Quercus}\footnote{Quercus
footnote}, did not support a number of language features. The full list of PHP
systems for that we evaluate our symbolic execution engine along with
descriptive code metrics can be found in Table \ref{corpus}.

@@

We evaluated this heuristic manually with a sample of 400 string
literals randomly selected from our corpus of PHP systems (see Table
\ref{corpus}). We measured for our heuristic classifier a precision of 94 percent, and a
recall of 50 percent. This means that six percent of the string literals are
classified incorrectly as false positives. In turn, the classifier is highly
distinctive as 96 percent of the string literals are classified correctly. The
recall of 50 percent means that half of the string literals, which we manually 
classified as output candidates, actually were responsive to the classifier. We
attempted to increase recall by looking for further distinctive properties to build a
classifier from, but could not do so without decreasing precision. 
In spite of missing half of the output candidates, we decided
to use a simple, yet precise and distinctive classifier. We rather work on a
smaller sample subset of actual output candidates with less false positives
than a larger set of string literals, where more are mistakenly classified as
output candidates.

@@

\paragraph{Dynamic Features}
According to a case study by Hills et al. \cite{Hills:2013:ESP:2483760.2483786}
of the feature usage in PHP, based on a corpus of state-of-the-art-projects,
dynamic includes are less frequently used than static includes, but usage
frequency is varying from system to system. Further work by Hills et al.
\cite{hills2014static,hills2014php} approaches static resolution of dynamic includes using
both context-insensitive resolution on file level by simplifying PHP constants; and context-sensitive on program level for transitive includes. In spite of
promising results, these approaches, similar to our results, face limitations
for truly dynamic includes if resolution is not sound due to information not
being available in the source code like database query results.

\paragraph{Input Generation for PHP}
\textcolor{gray}{Apollo \cite{artzi_finding_2008,artzi_finding_2010}, automated
test input generation using concolic testing \cite{DynamicWassermann}, test case generation
using search based software engineering \cite{Alshahwan2011}}
